Finding the True MVP

### 2024-12-03 paper/paper.qmd
Added Graphs and Rwender Test

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(knitr)
library(bayesplot)
library(StanHeaders)
library(Matrix)
library(lme4)
library(patchwork)

@@

```{r}
# 1986
NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1986[2, ]))
NBA1986 <- NBA1986[-c(1, 2), ]
colnames(NBA1986) <- new_col_names
NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)

@@

```{r}
Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")

@@

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

```{r}
# Fit a linear regression model
glm_model <- glm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)

@@

@@

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

@@

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(knitr)
library(bayesplot)
library(StanHeaders)
library(Matrix)
library(lme4)
library(patchwork)
library(kableExtra)

@@

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# 1986
NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1986[2, ]))
NBA1986 <- NBA1986[-c(1, 2), ]
colnames(NBA1986) <- new_col_names
NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)

@@

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")

@@

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

@@

@@

```{r}
# Fit a linear regression model
lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)

@@

next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
print(paste("Predicted MVP Index (mvp_table):", predicted_index_mvp_table))

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
print(paste("Predicted MVP Index (merged_real_mvp):", predicted_index_real_mvp))

@@

### Model Results
```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

# Filter MVPs from 2014 to 2024 for both datasets
mvp_table_filtered <- mvp_table %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

real_mvp_filtered <- merged_real_mvp %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

# Predict MVP index for 2025 and add it to both tables
next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_mvp_table <- NA
if (nrow(next_year_mvp_stats) > 0) {
  predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
}

mvp_table_filtered <- bind_rows(
  mvp_table_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_mvp_table
  )
)

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_real_mvp <- NA
if (nrow(next_year_real_mvp_stats) > 0) {
  predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
}

real_mvp_filtered <- bind_rows(
  real_mvp_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_real_mvp
  )
)

# Rename columns for clarity
mvp_table_filtered <- mvp_table_filtered %>%
  rename(
    `Simulated Data - Player` = Player,
    `Simulated Data - MVP Index` = MVP_index
  )

real_mvp_filtered <- real_mvp_filtered %>%
  rename(
    `Historical Data - Player` = Player,
    `Historical Data - MVP Index` = MVP_index
  )

# Join the two tables by Year
combined_table <- full_join(
  mvp_table_filtered,
  real_mvp_filtered,
  by = "Year"
)

# Create a styled table using kable
combined_table %>%
  kable(
    format = "html",
    caption = "Combined MVP Data (2014-2025)",
    col.names = c(
      "Year",
      "Simulated Data - Player",
      "Simulated Data - MVP Index",
      "Historical Data - Player",
      "Historical Data - MVP Index"
    )
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  add_header_above(c(" " = 1, "Simulated Data" = 2, "Historical Data" = 2))

@@

# Add a source column to the datasets
mvp_table <- mvp_table %>%
  mutate(Source = "Simulated MVPs")

merged_real_mvp <- merged_real_mvp %>%
  mutate(Source = "Historical MVPs")

# Combine datasets
combined_data <- bind_rows(
  mvp_table %>% select(Year, PTS, TRB, AST, Source),
  merged_real_mvp %>% select(Year, PTS, TRB, AST, Source)
)

# Plot for PTS
pts_plot <- ggplot(combined_data, aes(x = Year, y = PTS, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. PTS",
    x = "Year",
    y = "PTS"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for TRB
trb_plot <- ggplot(combined_data, aes(x = Year, y = TRB, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. TRB",
    x = "Year",
    y = "TRB"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for AST
ast_plot <- ggplot(combined_data, aes(x = Year, y = AST, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. AST",
    x = "Year",
    y = "AST"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Display plots separately
print(pts_plot)
print(trb_plot)
print(ast_plot)

@@

@@

@@
